---
output:
  pdf_document: default
  html_document: default
---
# Intro to Data Science - HW 7
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Hrishikesh Telang
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 2. I did this homework with help from the book and the professor and these Internet sources: 
#R documentation and StackOverflow
```

Last assignment we explored **data visualization** in R using the **ggplot2** package. This homework continues to use ggplot, but this time, with maps.  In addition, we will merge datasets using the built-in **merge( )** function, which provides a similar capability to a **JOIN in SQL** (don't worry if you do not know SQL). Many analytical strategies require joining data from different sources based on a **“key”** – a field that two datasets have in common. 

## Step 1: Load the population data

A.	Read the following JSON file, https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json and store it in a variable called **pop**.

Examine the resulting pop dataframe and add comments explaining what each column contains.



```{r}
library(jsonlite) #Call jsonlite library
library(RCurl) #Call Rcurl library
pop <- fromJSON(getURL('https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json')) 
#Read the JSON file using getURL() and parsing it using fromJSON()
head(pop) #Read the first six rows of 'pop'
```

B.	Calculate the **average population** in the dataframe. Why is using mean() directly not working? Find a way to correct the data type of this variable so you can calculate the average (and then calculate the average)

Hint: use **str(pop)** or **glimpse(pop)** to help understand the dataframe


```{r}
str(pop) #Trying to understand tge dataframe and the types of each dataframe
pop$population <- as.numeric(pop$population) #Converting the population column to numeric
mean(pop$population) #Assessing the average of population
```

C.	What is the population of the smallest city in the dataframe? Which state is it in?


```{r}
pop[which.min(pop$population), ] #Returns the row with the smallest city.
#The smallest city is Panama City, with population of only 36,877. 
#It is located in the state of Florida.
```

## Step 2: Merge the population data with the state name data

D)	Read in the state name .csv file from the URL below into a dataframe named **abbr** (for "abbreviation") – make sure to use the read_csv() function from the tidyverse package: <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv



```{r}
library(readr) #Call readr library (to read csv files)
abbr <- read_csv('https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv') 
#read the csv file and store in abbr
head(abbr) #Check the first six rows of abbr
```

E)	To successfully merge the dataframe **pop** with the **abbr** dataframe, we need to identify a **column they have in common** which will serve as the **“key”** to merge on. One column both dataframes have is the **state column**. The only problem is the slight column name discrepancy – in **pop**, the column is called **“state”** and in **abbr** – **“State.”** These names need to be reconciled for the merge() function to work. Find a way to rename **abbr’s “State”** to **match** the **state column in pop**.    


```{r}
colnames(abbr)[1] <- 'state'
abbr #df is of the dimension 51 x 2
```

F)	Merge the two dataframes (using the **‘state’ column** from both dataframes), storing the resulting dataframe in **dfNew**.


```{r}
dfNew <- merge(pop,abbr, 
               all.x=TRUE,by.x="state",by.y="state") 
#Performs an inner join between 'pop' and 'abbr' by the 'state' column.
```

G)	Review the structure of **dfNew** and explain the columns (aka attributes) in that dataframe.


```{r}
head(dfNew) #df is of the dimension 1000 x 8
```

## Step 3: Visualize the data

H)	Plot points (on top of a map of the US) for **each city**. Have the **color** represent the **population**.


```{r}
library(ggplot2); #Call ggplot library
library(maps); #Call maps library
library(ggmap); #Call ggmap library
library(mapproj) #Call mapproj library
library(tidyverse) #Call tidyverse library
us <- map_data("state") 
#Easily turn data from the maps package in to a data frame suitable for plotting with ggplot2.
dfNew$state <- tolower(dfNew$state) #lowercase all the state values to achieve consistency
us <- us %>% arrange(order) #arrange the 'order' of the us dataframe in an ascending manner (optional)
map <- ggplot(dfNew, aes(map_id= state)) #plot the maps using dfNew
map <- map + geom_map(map=us) #load the us map
map <- map + geom_point(color="magenta", aes(x=longitude,y=latitude, color=population)) 
#Plot the points on the map with magenta and set as per population
map #Display the map
```

I)	Add a block comment that criticizes the resulting map. It’s not very good.


```{r}
#The map is not clear and no proper results can be obtained from the map. I do not know
#what is the population density in each state.It neither shows any distribution 
#nor any patterns. I cannot infer anything from this figure.
```

## Step 4: Group by State


J)	Use group_by and summarise to make a dataframe of state-by-state population. Store the result in **dfSimple**.


```{r}
library(dplyr) #Call dplyr
dfSimple <- dfNew %>% group_by(state) %>% summarise(population = sum(population)) 
#It tries to group by state using tidyverse pipeline concept, and then summarizing it by performing sum()
head(dfSimple) #df is of the dimension 51 x 2
```

K)	Name the most and least populous states in **dfSimple** and show the code you used to determine them.


```{r}
dfSimple[which.max(dfSimple$population),] #returns the row with the maximum population
dfSimple[which.min(dfSimple$population),] #returns the row with the minimum population
```

## Step 5: Create a map of the U.S.,  with the color of the state representing the state population

L) Make sure to expand the limits correctly and that you have used **coord_map** appropriately.


```{r}
sample <- merge(us,dfSimple, all.x=TRUE,by.x="region",by.y="state") 
#merging the us and dfSimple dataframes with region and state columns of the respective datasets
#sample #Display the dataframe
us <- map_data("state") 
#Easily turn data from the maps package in to a data frame suitable for plotting with ggplot2.
sample$region <- tolower(sample$region) 
#lowercase all the state values to achieve consistency while mapping
us <- us %>% arrange(order) 
#arrange the 'order' of the us dataframe in an ascending manner (optional)
map2 <- ggplot(sample, aes(map_id= region)) #plot the maps using sample
map2 <- map2 + geom_map(map=us) #load the us map
map2 <- map2 + geom_polygon(color="white", aes(x=long,y=lat, fill=population)) + 
  coord_map(projection = "mercator") 
#Plot the points on the map with magenta and set as per population
map2 #Display the map
```
