---
output:
  pdf_document: default
  html_document: default
---
# Intro to Data Science - HW 6
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Hrishikesh Telang
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

**This module:** **Data visualization** is important because many people can make sense of data more easily when it is presented in graphic form. As a data scientist, you will have to present complex data to decision makers in a form that makes the data interpretable for them. From your experience with Excel and other tools, you know that there are a variety of **common data visualizations** (e.g., pie charts). How many of them can you name? <br>

The most powerful tool for data visualization in R is called **ggplot**. Written by computer/data scientist **Hadley Wickham**, this **“graphics grammar”** tool builds visualizations in layers. This method provides immense flexibility, but takes a bit of practice to master.


## Step 1: Make a copy of the data

A.	Read the **who** dataset from this URL: https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv <br>
into a new dataframe called **tb**. 

Your new dataframe, tb, contains a so-called **multivariate time series**: a sequence of measurements on 23 Tuberculosis-related (TB) variables captured repeatedly over time (1980-2013). Familiarize yourself with the nature of the 23 variables by consulting the dataset’s codebook which can be found here: https://intro-datascience.s3.us-east-2.amazonaws.com/TB_data_dictionary_2021-02-06.csv.


```{r}
tb <- read.csv("https://intro-datascience.s3.us-east-2.amazonaws.com/who.csv") #reading the data from csv file.
head(tb) #Display tb on console.
View(tb) #Viewing the contents of tb.
```

B.	How often were these measurements taken (in other words, at what frequency were the variables measured)? Put your answer in a comment. 


```{r}
#The variables were measured as per the frequency of each year starting from 1980 to 2008. The iso2 attribute denotes the country code in each year respectively.
```

## Step 2: Clean-up the NAs and create a subset

A.	Let's clean up the iso2 attribute in **tb** 

Hint: use *is.na()* -- well use *! is.na()* 


```{r}
tb<- tb[!is.na(tb$iso2),] #Filters out only those rows which are not NA values in iso2.
head(tb) #Display tb.
```

B. Create a subset of **tb** containing **only the records for Canada ("CA" in the iso2 variable)**. Save it in a new dataframe called **tbCan**. Make sure this new df has **29 observations and 23 variables**.


```{r}
tbCan <- subset(tb, tb$iso2 == 'CA') #subsets records only for iso2 variables which correspond to Canada.
tbCan #Display tbCan on console.
View(tbCan) #Viewing the contents of tbCan.
```

C. A simple method for dealing with small amounts of **missing data** in a numeric variable is to **substitute the mean of the variable in place of each missing datum**. <br>
This expression locates (and reports to the console) all the missing data elements in the variable measuring the **number of positive pulmonary smear tests for male children 0-4 years old** (there are 26 data points missing)


```{r}
tbCan$new_sp_m04[is.na(tbCan$new_sp_m04)]
```


    Error in eval(expr, envir, enclos): object 'tbCan' not found
    Traceback:



D.	Write a comment describing how that statement works.


```{r}
#The expression catches all the NA expressions with is.na() and returns a TRUE/FALSE vector. The outer expression ie: tbCan$new_sp_m04[,] essentially only returns a vector of those values from the attribute 'new_sp_m04' which were TRUE. This vector thus displays a vector of 26 NA values themselves.
```

E.	Write 4 more statements to check if there is missing data for the number of positive pulmonary smear tests for: **male and female** children 0-14 years old (**new_sp_m014** and **new_sp_f014**), and **male and female citizens 65 years of age and older**, respectively. What does empty output suggest about the number of missing observations?


```{r}
tbCan$new_sp_m014[is.na(tbCan$new_sp_m014)]
tbCan$new_sp_f014[is.na(tbCan$new_sp_f014)]
tbCan$new_sp_m65[is.na(tbCan$new_sp_m65)]
tbCan$new_sp_f65[is.na(tbCan$new_sp_f65)]

#The empty output suggests that there are no NA values present in all four of the attributes ie: new_sp_m014, new_sp_f014, new_sp_m65, and new_sp_f65.
```

There is an R package called **imputeTS** specifically designed to repair missing values in time series data. We will use this instead of mean substitution.  <br>
The **na_interpolation()** function in this package takes advantage of a unique characteristic of time series data: **neighboring points in time can be used to “guess” about a missing value in between**.

F.	Install the **imputeTS** package (if needed) and use **na_interpolation( )** on the variable from part C. Don’t forget that you need to save the results back to the **tbCan** dataframe. Also update any attribute discussed in part E (if needed).


```{r}
#install.packages('imputeTS') #This installs imputeTS
library(imputeTS) #We are calling the library of imputeTS
tbCan$new_sp_m04 <- na_interpolation(tbCan$new_sp_m04) #guesses the missing values in between
head(tbCan)
```

G.	Rerun the code from C and E above to check that all missing data have been fixed. 


```{r}
tbCan$new_sp_m014 <- na_interpolation(tbCan$new_sp_m014)
tbCan$new_sp_f014 <- na_interpolation(tbCan$new_sp_f014)
tbCan$new_sp_m65 <- na_interpolation(tbCan$new_sp_m65)
tbCan$new_sp_f65 <- na_interpolation(tbCan$new_sp_f65)

head(tbCan) #Displays tbCan
```

## Step 3: Use ggplot to explore the distribution of each variable
**Don’t forget to install and library the ggplot2 package.** Then: <br>
H.	Create a histogram for **new_sp_m014**. Be sure to add a title and briefly describe what the histogram means in a comment.


```{r}
library(ggplot2) #Calls the ggplot2 library function
ggplot(tbCan) + ggtitle('Frequency of positive pulmonary smear tests for male children 0-14 years old') + geom_histogram(aes(x=new_sp_m014), bins=10, fill='gray', col='black')

#The following histogram function here illustrates the positive pulmonary smear tests for male children between 0 to 14 years old. The highest frequency can be observed in children with pulmonary tests around 1-2, and lowest after 10 and around 6-7 tests.
```

I.	Create histograms (using ggplot) of each of the other three variables from E with ggplot( ). <br>
Which parameter do you need to adjust to make the other histograms look right?


```{r}
ggplot(tbCan) + ggtitle('Frequency of positive pulmonary smear tests for male children 0-14 years old') + geom_histogram(aes(x=new_sp_f014), bins=10, fill='red', col='black') 

ggplot(tbCan) + ggtitle('Frequency of positive pulmonary smear tests for male senior citizens over 65 years old') + geom_histogram(aes(x=new_sp_m65), bins=10, fill='green', col='black', binwidth=10)

ggplot(tbCan) + ggtitle('Frequency of positive pulmonary smear tests for female senior citizens over 65 years old') + geom_histogram(aes(x=new_sp_f65), bins=10, fill='blue', col='black', binwidth=5)

#I used the 'binwidth' parameter to make the necessary adjustments so as to make each of these histograms look right
```

## Step 4: Explore how the data changes over time
J.	These data were collected in a period of several decades (1980-2013). You can thus observe changes over time with the help of a line chart. Create a **line chart**, with **year** on the X-axis and **new_sp_m014** on the Y-axis.


```{r}
ggplot(tbCan) + geom_line(aes(x=year, y=new_sp_m014), colour='red')
```

K.	Next, create similar graphs for each of the other three variables. Change the **color** of the line plots (any color you want).


```{r}
ggplot(tbCan) + geom_line(aes(x=year, y=new_sp_f014), colour='cyan')
ggplot(tbCan) + geom_line(aes(x=year, y=new_sp_m65), colour='magenta')
ggplot(tbCan) + geom_line(aes(x=year, y=new_sp_f65), colour='black')
```

L.	Using vector math, create a new variable by combining the numbers from **new_sp_m014** and **new_sp_f014**. Save the resulting vector as a new variable in the **tbCan** df called **new_sp_combined014**. This new variable represents the number of positive pulmonary smear tests for male AND female children between the ages of 0 and 14 years of age. Do the same for SP **tests among citizens 65 years of age and older** and save the resulting vector in the tbCan variable called **new_sp_combined65**.


```{r}
tbCan$new_sp_combined014 <- tbCan$new_sp_m014 + tbCan$new_sp_f014 #adds the values and stores in new_sp_combined014
tbCan$new_sp_combined65 <- tbCan$new_sp_m65 + tbCan$new_sp_f65 #adds the values and stores in new_sp_combined65
head(tbCan)
```

M.	Finally, create a **scatter plot**, showing **new_sp_combined014** on the x axis, **new_sp_combined65** on the y axis, and having the **color and size** of the point represent **year**.


```{r}
ggplot(tbCan) + geom_point(aes(x=new_sp_combined014, y=new_sp_combined65, colour=year, size=year))
```

N.	Interpret this visualization – what insight does it provide?


```{r}
#It can be perceived that the number of Tuberculosis cases have significantly reduced over the past two decades. The proportion of children suffering from TB to that of the senior citizens was much higher in the 1980s. This is in stark contrast in the 2000s with a record of an all-time low cases. (0 to 100 for the senior citizens and a meager 1-10 for children aged between 0 to 14). This is a positive improvement.
```
