---
output:
  pdf_document: default
  html_document: default
---
## Intro to Data Science - Lab 3

# IST687 Section M003
# Professor Anderson
# Enter your name here: Hrishikesh Telang

#Select one of the below and add needed information
# 1. I did this homework by myself, with help from the book and the professor.

#Instructions:	Be	prepared	to	report the	whole	class at	the	conclusion	of	the	breakout	\
#group. Run	the	necessary	code	on	your own	instance	of	R-Studio.	Save	the	code:	It	
#will	be	useful	on	your	homework!

#1. Get	an	explanation	of	the	contents	of	the state.x77 data	set: help("state.x77")
```{r}
help("state.x77")
```

#2. Create	a	dataframe	from the	built-in	state.x77 data	set,	store	in	a	variable	named ‘dfStates77’
```{r}
dfStates77 <- state.x77
dfStates77
```

#3. Summarize	the	variables	in	your	dfStates77	data	set - using	the	summary()	function
```{r}
summary(dfStates77)
```

#4. Calculate	the	total	population	of	the	U.S. by adding	together	the	populations	of	
#each	of	the	individual	states in	dfStates77. Store	the	result	in	a	new	variable	
#called	totalPop77.
```{r}
totalPop77 <- sum(dfStates77[, 'Population'])
totalPop77
```

#5. Use	R	code	to	read	a	CSV	data	file	directly	from	the	web.	Store	the	dataset	
#into	a	new	dataframe,	called	dfStates17. The	URL	is:	
  " https://intro-datascience.s3.us-east-2.amazonaws.com/statesNew.csv"
#Note:	Use	the	function read_csv(	) to	read	in	the	data.	
#You	will	need	to	run	library(readr) or	library(tidyverse) before	you	can	run	
#read_csv(	). If	that	generates	an	error,	then	you	first	need	to	do	
#install.packages("readr") or install.packages("tidyverse")
```{r}
library(tidyverse)
dfStates17 <- read_csv('https://intro-datascience.s3.us-east-2.amazonaws.com/statesNew.csv')
dfStates17 <- data.frame(dfStates17)
```

#6. Summarize	the	variables	in	your	new	data	set,	using	the	summary	command.
```{r}
summary(dfStates17)
```

#7. The	data	you	now	have	stored	in	dfStates17 were	collected	in	2017.	As	such,	
#about	40	years	passed	between	the	two	data	collections.	Calculate	the	total	2017	
#population	of	the	U.S.	in	dfStates17 by	adding	together	the	populations	of	each	of	
#the	individual	states. Store	the	result	in	a	new	variable	called	totalPop17.	
```{r}
totalPop17 <- sum(dfStates17[,'population'])
totalPop17
```

#8. Create and	interpret a	ratio	of	totalPop77 to	totalPop17. Check	to	ensure	that	the	
#result	makes	sense!
```{r}
ratio <- totalPop77/totalPop17
ratio
```

#End	of	first	breakout
  
#Create	a	function	that,	given	population	and	area,	calculates	population	density	by	dividing	a	population	value	by	an	area	value.	Here	is	the	core	of	the	function:
```{r}
popDensity <- function (pop, area) {
  # Add your code below here:
  # Next, divide pop by area and store the result in a
  # variable called popDens
  popDens <- pop/area
      
  return(popDens) # This provides the function’s output
}
```

#9. After	you	finish	your	function,	make	sure	to	run	all	of the	lines	of	code	in	it	so	that	the	function	becomes	known	to	R.


#10. Make	a	fresh	copy	of	state.x77	into	dfStates77
```{r}
dfStates77 <- data.frame(state.x77)
dfStates77
```

#11. Store	the	population	vector	in	a	variable	called	tempPop.	Adjust	the	tempPop	as	needed	(based	on	your	analysis	at	the	end	of	the	first	breakout)
```{r}
tempPop <- dfStates77[,'Population'] * 1000
tempPop
```

#12. Store	the	area	vector	in	a	variable,	called	tempArea	
```{r}
tempArea <- dfStates77[,'Area']
tempArea
```

#13. Now	use	tempPop	and	tempArea	to	call	your	function:popDensity(tempPop,	tempArea)
```{r}
popDensity(tempPop,	tempArea)
```


#14. Store	the	results	from the	previous	task	in	a	column	of	the	dfStates77	dataframe,	called	popDensity.
```{r}
dfStates77$popDensity <- popDensity(tempPop,	tempArea)
dfStates77
```

#15. Use	which.max(	)	and	which.min(	)	to	reveal	which	is	the	most	densely	populated	and	which	is	the	least	densely	populated	state.
#Make	sure	that	you	understand	the	number	that	is	revealed	as	well	as	the	name	of	the	state.
```{r}
dfStates77[which.max(dfStates77$popDensity),]
dfStates77[which.min(dfStates77$popDensity),]
```


#16. Using	tidyverse,	sort	the	dataframe	using	the	popDensity	attribute,	then	using	
#the	slice	function,	show	the	first	row	in	the	sorted	database.	
```{r}
library('tidyverse')
sortedDF <- dfStates77 %>% arrange(popDensity)
sortedDF <- slice(sortedDF, 1)
sortedDF
```


#17. How	was	the	dataframe	sorted	(was	the	minimum	first	or	the	maximum)?
```{r}
#It was the minimum first.
```

