---
output:
  pdf_document: default
  html_document: default
---
# Intro to Data Science - HW 10

##### Copyright 2021, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Hrishikesh Telang
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

**Association mining** can be applied to many data problems beyond the well-known example of **finding relationships between different products in customer shopping data**. In this homework assignment, we will explore **real data** from the banking sector and look for **patterns associated with the likelihood of responding positively to a direct marketing campaign and signing up for a term deposit with the bank (stored in the variable “y”)**. <br>
You can find out more about the variables in this dataset here: https://archive.ics.uci.edu/ml/datasets/bank+marketing

## Part 1: Explore Data Set

A.	Read the contents of the following URL to a dataframe called **bank** <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv

**Hint**: Even though this is a .csv file, chances are R won’t be able to read it in correctly using the read_csv() function. If you take a closer look at the contents of the URL file, you may notice each field is separated by a **semicolon** (;) rather than a comma. 

In situations like this, consider using either read.csv or read.table, with two additional parameters. *sep=";"* defines how the data is seperated (the default is a comma), and *header=TRUE* defines that there is a header line in the dataset. 


```{r}
library(readr) #Initiate readr library
bank <- read.csv('https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv', 
                 sep=";", header=TRUE)
#read csv file from the internet
head(bank) #Show the head of the bank
tail(bank) #Show the tail of the bank
```

Make sure there are **41,188** rows and **21** columns in your **bank** df.


```{r}
#To cross check that, we have used str(), nrow() and ncol()
str(bank) #returns the structure (datatypes) of the bank dataset
nrow(bank) #returns number of rows
ncol(bank) #returns number of columns
```

B.	Next, we will focus on some key factor variables from the dataset, and convert a few numeric ones to factor variables. Execute the following command.  Write a comment describing how the conversion for each numeric variable works and what are the variables in the resulting dataframe.


```{r}
bank_new <- data.frame(job=as.factor(bank$job),
                     marital=as.factor(bank$marital),
                     housing_loan=as.factor(bank$housing),
                     young=as.factor(bank$age<median(bank$age)),
                     contacted_more_than_once=as.factor(bank$campaign>1),
                     contacted_before_this_campaign=as.factor(bank$previous<0),
                     success=as.factor(bank$y))
```

C.	Count the number of successful term deposit sign-ups, using the table( ) command on the **success** variable.


```{r}
table(bank_new$success) #returns the number of rows that have "no" and "yes" labels
```

D.	Express the results of problem C as percentages by sending the results of the table( ) command into the prop.table( ) command.


```{r}
prop.table(table(bank_new$success)) #returns the percentage of the number of rows
#with "no" and "yes" labels
```

E.	Using the same techniques, show the percentages for the **marital** and **housing_loan** variables as well.


```{r}
print("%%%%marital%%%%")
table(bank_new$marital) #returns the number of rows that have "no" and "yes" labels
prop.table(table(bank_new$marital))
#returns the percentage of the number of rows with "no" and "yes" labels

print("%%%%housing_loan%%%%")
table(bank_new$housing_loan) #returns the number of rows that have "no" and "yes" labels
prop.table(table(bank_new$housing_loan))
#returns the percentage of the number of rows with "no" and "yes" labels
```

## Part 2: Coerce the data frame into transactions

F.	Install and library two packages: **arules** and **arulesViz**.


```{r}
#install.packages('arules')
#install.packages('arulesViz')
library(arules)
library(arulesViz)
```

G.	Coerce the **bank_new** dataframe into a **sparse transactions matrix** called **bankX**.


```{r}
bankX <- as(bank_new, "transactions")
bankX
```

H.	Use the itemFrequency( ) and itemFrequencyPlot( ) commands to explore the contents of **bankX**. What do you see?


```{r}
itemFrequency(bankX) #Shows item frequency for each categorical value
itemFrequencyPlot(bankX) #Plots the frequency distribution

# I see a frequency graph of all the occurences of attributes from the sparse 
#transactions matrix
#The Y axis signifies the frequency of the number of times the item has been selected.
#(In other words, Y axis indicates the support.)
#Based on this observation, attributes such as contacted_before_this_campaign has 
#had the highest occurence, followed by success, then marital and so forth.
```

I.	This is a fairly large dataset, so we will explore only the first 10 observations in the **bankX** transaction matrix: 


```{r}
inspect(bankX[1:10]) 
# With this code, we are exploring only the first 10 observations in the 'bankX' transaction matrix
```

Explain the difference between **bank_new** and **bankX** in a block comment:

```{r}
#bank_new is the new dataset that is returned after all the variables which looked
#categorical in nature are converted into factor variables, whereas, bankX is the
#a sparse transaction matrix that results from the categorical valuables of bank_new.
#It contains the frequency values as a probability distribution of all the 
#different columns along with their categorical values. 
```

## Part 3: Use arules to discover patterns

**Support** is the proportion of times that a particular set of items occurs relative to the whole dataset. <br>
**Confidence** is proportion of times that the consequent occurs when the antecedent is present. <br>

J.	Use **apriori** to generate a set of rules with support over 0.005 and confidence over 0.3, and trying to predict who successfully signed up for a term deposit. <br>
**Hint:** You need to define the **right-hand side rule (rhs)**.


```{r}
#I have listed the supp=0.005 and conf= 0.3 to satisfy the question of the prompt.
#I also used success=yes to predict who successfully signed up for a term deposit.
rules3 <- apriori(bank_new, parameter = list(supp=0.005, conf= 0.3), 
                  appearance=list(default="lhs", rhs="success=yes"), 
                  control=list(verbose=F))
```

K.	Use inspect() to review of the **ruleset**. 


```{r}
inspect(rules3)
```

L.	Use the output of inspect( ) or inspectDT( ) and describe **any 2 rules** the algorithm found.  


```{r}
#The rule {job=student, marital=single, young=TRUE, contacted_before_this_campaign=FALSE} 
#=> {success=yes} shows a lift of 2.870582. This is the highest value computed for the given
#rule in the ruleset. This is followed by two second highest rules {job=student, marital=single} 
#=> {success=yes} and {job=student, marital=single, contacted_before_this_campaign=FALSE} 
#=> {success=yes} with equal lifts, confidences and supports and will be treated 
#equally based on the conditions set at training the apriori model.
```
