---
output:
  pdf_document: default
  html_document: default
---
# Intro to Data Science - HW 5

##### Copyright 2021, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: Hrishikesh Telang
```

### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).


```{r}
library(RCurl) #We are calling the Rcurl package
library(jsonlite) #We are calling the jsonlite package
dataset <- getURL("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json") #We are calling the JSON file and storing in dataset
readlines <- jsonlite::fromJSON(dataset) #This function converts JSON data to an R object
df <- readlines$objects$person #This line stores all the JSON data into df
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).


```{r}
head(df)
tail(df)
View(df)
```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>

```{r}
#1. The dataset contains the information of all the US Senators
#2. There are 100 rows and every row represents a senator
#3. There are 17 columns and every column represents the unique ID, attributes (or personal information of each senator)
```

## Part 2: Investigate the resulting dataframe 

C.	How many senators are women? 


```{r}
#Using nrow
nrow(df[df$gender == 'female',])

#Using sum
sum(df$gender == 'female')
```

D.	How many senators have a YouTube account?  


```{r}
#Using sum
sum(is.na(df$youtubeid) == FALSE)

#Using nrow
nrow(df[is.na(df$youtubeid) == FALSE,])
```

E.	How many women senators have a YouTube account?


```{r}
nrow(df[df$gender == 'female' & is.na(df$youtubeid) == FALSE,])
```

F.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   


```{r}
youtubeWomen <- df[df$gender == 'female' & is.na(df$youtubeid) == FALSE,]
View(youtubeWomen)
```

G.	What does running this line of code do? Explain in a comment: 


```{r}
youtubeWomen$year <- substr(youtubeWomen$birthday,1,4)
#The code of line essentially subsets the year from the 'birthday' column and adds it in a new 'year' column
View(youtubeWomen)
```

H.	Use this new variable to calculate the mean **birthyear** in **youtubeWomen**. <br>
**Hint:** You may need to convert it to numeric first.


```{r}
youtubeWomen$year <- as.numeric(as.character(youtubeWomen$year))
mean(youtubeWomen$year)
```

I.	Make a histogram of the **birthyears** of senators in **youtubeWomen**. Add a comment describing the shape of the distribution.


```{r}
hist(youtubeWomen$year,
     main='Distribution of the age of the Female Senators who are on YouTube',
     xlab='Birth Years')

#The shape of the bell curve is slightly shifted towards the left, which means that there is a greater frequency of older female senators
```
